---
title: 'Example MR analysis'
author: "Austin Hammermeister Suger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())

## Load R packages ##

## Install pacman from CRAN
install.packages("pacman")
library(pacman)

## Install and load tidyverse and kableExtra from CRAN
p_load(tidyverse,kableExtra)

# ## Install and load TwoSampleMR
install.packages(
  'TwoSampleMR',
  repos = c(
    'https://mrcieu.r-universe.dev/bin/linux/noble-x86_64/4.5/',
    'https://p3m.dev/cran/__linux__/noble/latest',
    'https://cloud.r-project.org'
  )
)

library(TwoSampleMR)
```


# Load harmonized MR data
- We already have a prepared MR instrument containing 363 SNPs associated with height
- The GWAS summary statistics for height came from PMID 36224396
- The GWAS summary statistics for systolic blood pressure came from PMID 34594039 
- Both datasets are from analyses of individuals of East Asian genetic ancestry 

```{r load MR data}
MR_data = readRDS("~/COURSE_MATERIALS/MR_example/Example_harmonized_MR_data.rds")
```

# Inspect the MR data
- What type of information is contained in this dataset?
- Why is it required for MR?

```{r view MR data}
glimpse(MR_data) 
```

# Run an MR analysis with standard MR methods

```{r run basic MR }
basic_MR_results = TwoSampleMR::mr(MR_data)
```


# Produce a table and scatterplot for the results

```{r MR results table and scatterplot}
basic_MR_results %>% 
  select(-c(id.exposure,id.outcome)) %>%
  kable() %>%
  kable_classic_2()
TwoSampleMR::mr_scatter_plot(basic_MR_results,MR_data)
```

# Run an MR-Egger test for pleiotropy (MR-Egger intercept)

```{r run MR-Egger }
MR_pleiotropy_results = TwoSampleMR::mr_pleiotropy_test(MR_data)
MR_pleiotropy_results %>% 
  select(-c(id.exposure,id.outcome)) %>%
  kable() %>%
  kable_classic_2()
```


# Optional: test the association with other MR methods
- If you have time, you can look at running MR analyyses using other MR methods
- The mr_method_list() function provides a list of other MR methods included in the package

```{r run other MR methods }
TwoSampleMR::mr_method_list()  %>%
  select(c(name,PubmedID)) %>%
  kable() %>%
  kable_classic_2()
```

# Additional information

More detailed information on how to use the TwoSampleMR R package to perform MR analyses can be found here: https://mrcieu.github.io/TwoSampleMR/

